---
title: "Generating Pedigrees for Trimmed Intermountain, Utah Slope Pedigrees, < -5"
subtitle: "Report to: Marcus Pezzolesi"
author: Scott Frodsham
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: margin=1cm
output:
  pdf_document:
    latex_engine: xelatex
    highlight: "default"
mainfont: Arial
monofont: Monaco
fontsize: 11pt
toc: yes
---

# Introduction

We have large pedigrees that have anuy diabetic and slow decliner that is related to a rapid decliner.

```{r load ped file}
library(tidyverse)

families <- read_csv("~/Documents/Projects/eGFR/EDW_IHC_RFD_Pedigrees/20190717DavidMarcusM5Pedigree/mappingDiabetics/all_secondCousinDiabeticDecliner_peds.csv", col_names = c("kindredID", "egoUPDBID", "paUPDBID", "maUPDBID", "sex", "deceased", "updb_id", "dist_id"))
```

We attempt to make a pedigree object again:

```{r make bigped 3}
pedList <- pedigree(famid = families$kindredID, id = families$egoUPDBID, dad = families$paUPDBID, mom = families$maUPDBID, sex = families$sex, status = families$deceased)
#write_csv(families, "~/Documents/Projects/eGFR/EDW_IHC_RFD_Pedigrees/20190717DavidMarcusM5Pedigree/mappingDiabetics/kinship2_ready_pedigree_dm.csv")

famNumList <- unique(families$kindredID)

#plotGenericPeds <- function(x) {
#  pedigree <- pedList[as.character(x)]
#  pdf(paste0("/Users/u6013142/Documents/Projects/eGFR/EDW_IHC_RFD_Pedigrees/20190717DavidMarcusM5Pedigree/mappingDiabetics/allSeparatePedPlots/", x, ".pdf"), width = 24, height = 12)
#  plot.pedigree(pedigree, id = rep("", length(pedigree$famid), cex = 0.4, symbolsize = 1))
#  mtext(paste0("Family ", x), line = 2, adj = 0, cex = 1.5)
#  dev.off()
#}
#
#lapply(famNumList, plotGenericPeds)
```

It worked! Before drawing the pedigrees, we want to add bring in some clinical information:

```{r import clinical data, message=FALSE, error=FALSE}
slopes <- read_csv("/Users/u6013142/Documents/Projects/eGFR/EDW_IHC_RFD_Pedigrees/medical_data/all_uddb_slopes.csv")
diags <- read_csv("/Users/u6013142/Documents/Projects/eGFR/EDW_IHC_RFD_Pedigrees/medical_data/all_uddb_diagnoses.csv")
egfr <- read_csv("/Users/u6013142/Documents/Projects/eGFR/EDW_IHC_RFD_Pedigrees/medical_data/uddb_egfrs_102219.csv")
sum(unique(slopes$Patient_ID) %in% unique(egfr$Patient_ID))
acr <- read_csv("/Users/u6013142/Documents/Projects/eGFR/EDW_IHC_RFD_Pedigrees/medical_data/uddb_acrs.csv")
acr <- acr %>% filter(!is.na(OBS_VALUE))
acr$RESULT_DTM <- as.Date(acr$RESULT_DTM, format = "%m/%d/%Y")
patientIDs <- read_csv("/Users/u6013142/Documents/Projects/eGFR/EDW_IHC_RFD_Pedigrees/medical_data/uddb_patientIDs.csv", col_types = "cicc")
# used to keep duplicates with this method but we're just going to delete them
#collapsed_patientIDs <- patientIDs %>% group_by(Patient_ID) %>% fill(DIST_ID, UPDB_ID) %>% fill(DIST_ID, UPDB_ID, .direction = 'up') %>% distinct()
#collapsed_patientIDs <- collapsed_patientIDs %>% distinct(Patient_ID, .keep_all = TRUE)
```

We make the patientIDs dataframe useful, getting rid of the `DIST_` and `UPDB_` prefixes also remove `NA` suffixes, we remove duplicates:

```{r process patientIDs}
#patientIDs_patid <- as_tibble(str_split(patientIDs$Patient_ID, "N", simplify = TRUE))
#names(patientIDs_patid) <- c("id", "a")
patientIDs_updb <- as_tibble(str_split(patientIDs$UPDB_ID, "_", simplify = TRUE))
names(patientIDs_updb) <- c("updb", "id")
patientIDs_dist <- as_tibble(str_split(patientIDs$DIST_ID, "_", simplify = TRUE))
names(patientIDs_dist) <- c("dist", "id")

patientIDs$UPDB_ID <- ifelse(is.na(patientIDs_updb$updb), NA, patientIDs_updb$id)
patientIDs$DIST_ID <- ifelse(is.na(patientIDs_dist$dist), NA, patientIDs_dist$id)
#patientIDs$Patient_ID <- patientIDs_patid$id

patientID_dups <- patientIDs$Patient_ID[duplicated(patientIDs$Patient_ID)]
patientIDs <- patientIDs %>% filter(!Patient_ID %in% patientID_dups)
```

We need to make a table that indicates the first and last eGFR for each patient:

```{r first last egfr}
egfr_last <- egfr %>% group_by(Patient_ID) %>% summarize(last_egfr = EGFR[which.max(RESULT_DTM)])
acr_last <- acr %>% group_by(Patient_ID) %>% summarize(last_acr = OBS_VALUE[which.max(RESULT_DTM)])
```

We join these values into one table:

```{r remove prefixes}
diags <- left_join(diags, patientIDs[ , c(1,2)])
slopes <- left_join(slopes, patientIDs[ , c(1,2)])
egfr_last_join <- left_join(egfr_last, patientIDs[ , c(1,2)])
acr_last <- left_join(acr_last, patientIDs[ , c(1,2)])
```

We need to filter some of these diagnoses down so that we're only dealing with the ones we need. First we process diabetes type:

```{r process dm}
dmType <- read_csv("~/Documents/Projects/eGFR/EDW_IHC_RFD_Pedigrees/20190717DavidMarcusM5Pedigree/mappingDiabetics/uddb_data/uddb_ids_dmTypes.csv")

dmType_patid <- as_tibble(str_split(dmType$Patient_ID, "N", simplify = TRUE))
names(dmType_patid) <- c("id", "a")
dmType_updb <- as_tibble(str_split(dmType$UPDB_ID, "_", simplify = TRUE))
names(dmType_updb) <- c("updb", "id")
dmType_dist <- as_tibble(str_split(dmType$DIST_ID, "_", simplify = TRUE))
names(dmType_dist) <- c("dist", "id")

dmType$UPDB_ID <- ifelse(is.na(dmType_updb$updb), NA, dmType_updb$id)
dmType$DIST_ID <- ifelse(is.na(dmType_dist$dist), NA, dmType_dist$id)
dmType$Patient_ID <- dmType_patid$id

dmType_dups <- dmType$Patient_ID[duplicated(dmType$Patient_ID)]
dmType <- dmType %>% filter(!Patient_ID %in% dmType_dups)

dmType$CleanFinalType <- ifelse(grepl("T2D", dmType$FinalType) | grepl("T3D", dmType$FinalType), 2, 1)
rm(patientIDs_dist, patientIDs_patid, patientIDs_updb, dmType_dist, dmType_patid, dmType_updb, patientID_dups, dmType_dups)
```

Now the rest of the diagnoses:

```{r filter diags}
# with aki
#esrd_diags <- diags %>% filter(ICD_CODE %in% c("584.9", "N17.9", "Z99.2", "V45.11", "585.6", "N18.6", "V42.0", "Z94.0"))
# without aki
esrd_diags <- diags %>% filter(ICD_CODE %in% c("Z99.2", "V45.11", "585.6", "N18.6", "V42.0", "Z94.0"))
esrd_list <- unique(esrd_diags$Patient_ID)

esrd_below_15 <- slopes %>% filter(CutoffEvent == "BELOW_15")
esrd_below_15_list <- unique(esrd_below_15$Patient_ID)

esrd_list <- unique(c(as.character(esrd_list), as.character(esrd_below_15_list)))

dmType$esrd <- ifelse(dmType$Patient_ID %in% esrd_list, "ESRD", "")  
slope_esrd_diabetes <- left_join(dmType, slopes[ , c(1,4)])

slope_esrd_diabetes_lastacr <- left_join(slope_esrd_diabetes, acr_last)
slope_esrd_dm_lastacr_lastegfr <- left_join(slope_esrd_diabetes_lastacr, egfr_last)

families$egoUPDBID <- as.character(families$egoUPDBID)
families_phenos <- left_join(families, slope_esrd_dm_lastacr_lastegfr, by = c("egoUPDBID" = "Patient_ID"))
families_phenos$Slope[families_phenos$Slope == "Rejected"] <- NA
families_phenos$Slope <- as.numeric(families_phenos$Slope)

families_phenos$rapidDecliner <- ifelse(families_phenos$Slope <= -5, 1, 0)
# add esrd affection line here
families_phenos$updb_id <- ifelse(is.na(families_phenos$updb_id), families_phenos$UPDB_ID, families_phenos$updb_id)
families_phenos$dist_id <- ifelse(is.na(families_phenos$dist_id), families_phenos$DIST_ID, families_phenos$dist_id)

write_csv(families_phenos, "~/Documents/Projects/eGFR/EDW_IHC_RFD_Pedigrees/20190717DavidMarcusM5Pedigree/mappingDiabetics/460_full_secondCousinDiabeticDecliner_pyTrimmed_peds_kinship_pheno_ready.csv")
```

Now we draw the pedigrees:

```{r draw pedigrees}
# make a list of the families
families_phenos <- read_csv("~/Documents/Projects/eGFR/EDW_IHC_RFD_Pedigrees/20190717DavidMarcusM5Pedigree/mappingDiabetics/460_full_secondCousinDiabeticDecliner_pyTrimmed_peds_kinship_pheno_ready.csv")

famList <- unique(as.character(families_phenos$kindredID))

pedList <- pedigree(famid = families_phenos$kindredID, id = families_phenos$egoUPDBID, dad = families_phenos$paUPDBID, mom = families_phenos$maUPDBID, sex = families_phenos$sex, status = families_phenos$deceased, affected = families_phenos$rapidDecliner, missid = 0)

plotSepPedsColor <- function(x) {
  #x <- 584
  fullInfo <- families_phenos %>% filter(kindredID == x)
  print(x)
  fullInfo$color <- ifelse(is.na(fullInfo$updb_id) & is.na(fullInfo$dist_id), 1,
                          ifelse(!is.na(fullInfo$updb_id) & !is.na(fullInfo$dist_id), 3, 
                          ifelse(!is.na(fullInfo$updb_id) & is.na(fullInfo$dist_id), 4,
                          ifelse(is.na(fullInfo$updb_id) & !is.na(fullInfo$dist_id), 2,5))))
                          #ifelse(fullInfo$dist_id %in% ngs_overlap$dist_id, 5, 6)))))
  pedigree <- pedList[as.character(x)]
  pdf(paste0("/Users/u6013142/Documents/Projects/eGFR/EDW_IHC_RFD_Pedigrees/20190717DavidMarcusM5Pedigree/mappingDiabetics/allSeparatePedPlots/", x, ".pdf"), width = 24, height = 12)
  #listOfIdsDM <- ifelse(is.na(fullInfo$CleanFinalType), "", fullInfo$egoUPDBID)
  listOfIds <- ifelse(is.na(fullInfo$Slope) & is.na(fullInfo$CleanFinalType), 
                      "", 
                      ifelse(is.na(fullInfo$Slope) & !is.na(fullInfo$CleanFinalType),
                             paste0(fullInfo$updb_id, "\n", fullInfo$dist_id, "\n", fullInfo$CleanFinalType, "\n", fullInfo$esrd), 
                             ifelse(!is.na(fullInfo$Slope) & !is.na(fullInfo$esrd), 
                                    paste0(fullInfo$updb_id, "\n", fullInfo$dist_id, "\n", fullInfo$CleanFinalType, "\n", fullInfo$last_egfr, "\n", fullInfo$Slope, "\n", fullInfo$last_acr, "\n", fullInfo$esrd), 
                                    paste0(fullInfo$updb_id, "\n", fullInfo$dist_id, "\n", fullInfo$CleanFinalType, "\n", fullInfo$last_egfr, "\n", fullInfo$Slope, "\n", fullInfo$last_acr))))
  #listOfIds <- ifelse(is.na(fullInfo$Slope), 
  #                    "", 
  #                    ifelse(!is.na(fullInfo$esrd), 
  #                           paste0(fullInfo$updb_id, "\n", fullInfo$dist_id, "\n", fullInfo$CleanFinalType, "\n", fullInfo$last_egfr, "\n", fullInfo$Slope, "\n", fullInfo$last_acr, "\n", fullInfo$esrd), 
  #                           paste0(fullInfo$updb_id, "\n", fullInfo$dist_id, "\n", fullInfo$CleanFinalType, "\n", fullInfo$last_egfr, "\n", fullInfo$Slope, "\n", fullInfo$last_acr)))
  fullInfo$rapidDecliner <- ifelse(!is.na(fullInfo$esrd), 1, fullInfo$rapidDecliner)
  plot.pedigree(pedigree, id = listOfIds, col = fullInfo$color, affected = fullInfo$rapidDecliner, cex = 0.5, symbolsize = 1.75)
  mtext(paste0("Family ", x), line = 2, adj = 0, cex = 1.5)
  dev.off()
}

#pdf(paste0("/Users/u6013142/Documents/Projects/eGFR/EDW_IHC_RFD_Pedigrees/20190717DavidMarcusM5Pedigree/mappingDiabetics/460_pedigreePlots.pdf"), width = 24, height = 12)
lapply(famList, plotSepPedsColor)
#dev.off()

# Only veryLegible families
veryLegibleFamIds <- scan("/Users/u6013142/Documents/Projects/eGFR/EDW_IHC_RFD_Pedigrees/20190717DavidMarcusM5Pedigree/mappingDiabetics/veryLegibleFamIds_slowDeclinersAndDM_phi0156.csv")
illegibleFamIds <- famList[!famList %in% veryLegibleFamIds]
write.table(illegibleFamIds, "/Users/u6013142/Documents/Projects/eGFR/EDW_IHC_RFD_Pedigrees/20190717DavidMarcusM5Pedigree/mappingDiabetics/illegibleFamIds_211_slowDeclinersAndDM_phi0156.csv", row.names = FALSE, col.names = FALSE, quote = FALSE)

pdf(paste0("/Users/u6013142/Documents/Projects/eGFR/EDW_IHC_RFD_Pedigrees/20190717DavidMarcusM5Pedigree/mappingDiabetics/249_veryLegiblePedigreePlots.pdf"), width = 24, height = 12)
lapply(veryLegibleFamIds, plotSepPedsColor)
dev.off()

pdf(paste0("/Users/u6013142/Documents/Projects/eGFR/EDW_IHC_RFD_Pedigrees/20190717DavidMarcusM5Pedigree/mappingDiabetics/211_illegiblePedigreePlots.pdf"), width = 24, height = 12)
lapply(illegibleFamIds, plotSepPedsColor)
dev.off()

#families_phenos_sum_5 <- families_phenos %>% group_by(kindredID) %>% summarize(n = n())
#
#plotSepPedsColorAllTogether <- function(x) {
#  fullInfo <- families_phenos %>% filter(kindredID == x)
#  fullInfo$color <- ifelse(is.na(fullInfo$updb_id) & is.na(fullInfo$dist_id), 1,
#                          ifelse(!is.na(fullInfo$updb_id) & !is.na(fullInfo$dist_id), 3, 
#                          ifelse(!is.na(fullInfo$updb_id) & is.na(fullInfo$dist_id), 4,
#                          ifelse(is.na(fullInfo$updb_id) & !is.na(fullInfo$dist_id), 2, 5))))
#  pedigree <- pedList[x]
#  plot.pedigree(pedigree, id = ifelse(is.na(fullInfo$UPDB_dist_id), "", paste0(fullInfo$updb_id, "\n", fullInfo$dist_id, "\n", fullInfo$TYPE_1OR2, "\n", fullInfo$last_egfr, "\n", fullInfo$Slope, "\n", fullInfo$last_acr, "\n", fullInfo$esrd)),
#                          col = fullInfo$color, affected = fullInfo$affection, cex = 0.45, symbolsize = 1)
#  mtext(paste0("Family ", x), line = 2, adj = 0, cex = 1.5)
#}
#
## this vvvvv needs to be run in a vanilla R script (not an Rmd)
#pdf(paste0("/Users/u6013142/Documents/Projects/eGFR/EDW_IHC_RFD_Pedigrees/20190717DavidMarcusM5Pedigree/pedigree_M5_trimmed.pdf"), width = 14, height = 6)
#lapply(famList, plotSepPedsColorAllTogether)
#dev.off()
```
