---
title: "drawing_full_high_risk_peds"
output: html_document
---

## Load packages and pedigree file, set working dir

```{r setup}
library(kinship2)
library(tidyverse)
library(knitr)
library(reticulate)

setwd('Y:/IRB_00096551/medalist_like_pedigrees_12_2021')

# Load pedigree file
families <- read_csv('20211202_T1D_LCA_Ped/David_Marcus_T1D_LCA_Trimmed_Pedigree.csv')
fullPed <- families
```

We change some of the encodings so that `kinship2` will accept them:

```{r change encoding}
families$sex[families$sex == "M"] <- "male"
families$sex[families$sex == "F"] <- "female"
families$sex[families$sex == "U"] <- "unknown"

families$affection[families$affection == 1] <- 0
families$affection[families$affection == 10] <- 1

families$deceased <- families$Dyr
families$deceased[!is.na(families$Dyr)] <- 1
families$deceased[is.na(families$Dyr)] <- 0

families$paUPDBID[is.na(families$paUPDBID)] <- 0
families$maUPDBID[is.na(families$maUPDBID)] <- 0
```

We will only keep the columns we need to create the pedigree:

```{r keep essential cols}
families <- families %>% select(kindredID, egoUPDBID, paUPDBID, maUPDBID, sex, deceased, affection)


#families <- families %>% filter(KindredID == 562503)
```

We attempt to make a pedigree object:

```{r make pedigree object}
#bigped <- pedigree(famid = families$kindredID, id = families$egoUPDBID, dad = families$paUPDBID, mom = families$maUPDBID, sex = families$sex, status = families$deceased, affected = families$affection)
```

So it looks like there are individuals who show up as parents, but not individuals (no egoUPDBID, etc),
lets create a new row for each of them. Lets save what we have as a tsv then open it in python.

```{r save file as tsv}
# if we try to save large numeric types scientific notation is introduced, so we change them to integers
families <- families %>% mutate_if(is.numeric, as.integer)
write_delim(families, "Z:/temp/medalist_like_ped.tsv", delim = "\t")
```

The following script is used to identify lists of fathers and mothers who do not have their 'own' line (in which their ID is in the `egoUPDBID` column):

```{python add missing parents}
with open('Z:/temp/medalist_like_ped.tsv') as full_fam_ped:
    full_fam_ped.readline()
    egoDict = {}
    momDict = {}
    dadDict = {}
    for line in full_fam_ped:
        line = line.strip().split("\t")
        family = int(line[0])
        ego = int(line[1])
        dad = int(line[2])
        mom = int(line[3])
        if ego != 0:
            if ego not in egoDict:
                egoDict[ego] = [family]
            else:
                egoDict[ego].append(family)
        if dad != 0:
            if dad not in dadDict:
                dadDict[dad] = [family]
            else:
                dadDict[dad].append(family)
        if mom != 0:
            if mom not in momDict:
                momDict[mom] = [family]
            else:
                momDict[mom].append(family)
    missingDads = []
    missingMoms = []
    for mom, fams in momDict.items():
        for fam in fams:
            if mom in egoDict:
                if fam not in egoDict[mom]:
                    missingMom = [mom, fam]
                    missingMoms.append(missingMom)
            else:
                missingMom = [mom, fam]
                missingMoms.append(missingMom)
        
    for dad, fams in dadDict.items():
        for fam in fams:
            if dad in egoDict:
                if fam not in egoDict[dad]:
                    missingDad = [dad, fam]
                    missingDads.append(missingDad)
            else:
                missingDad = [dad, fam]
                missingDads.append(missingDad)
```

We use the lists created by the python script to create new rows. We then append it to our `full_families` dataframe:

```{r create entries}
missingDads <- py$missingDads
missingMoms <- py$missingMoms
egoDict <- py$egoDict
momDict <- py$momDict
dadDict <- py$dadDict

missingParents <- function(x, y) {
  data.frame(
    kindredID = x[2],
    egoUPDBID = x[1],
    paUPDBID = 0,
    maUPDBID = 0,
    sex = y,
    deceased = 1,
    affection = NA,
    stringsAsFactors = FALSE)
}

if (length(missingDads) > 0) {
  dadRows <- lapply(missingDads, missingParents, "male")
  dadDF <- do.call(rbind, dadRows)
  dadsReady <- TRUE
} else {
  dadsReady <- FALSE
}

if (length(missingMoms) > 0) {
  momRows <- lapply(missingMoms, missingParents, "female")
  momDF <- do.call(rbind, momRows)
  momsReady <- TRUE
} else {
  momsReady <- FALSE
}

if (momsReady & dadsReady) {
  full_families_dads <- rbind(families, dadDF)
  full_families_moms_dads <- rbind(full_families_dads, momDF)
  families <- full_families_moms_dads %>% mutate_if(is.numeric, as.integer)
} else if (momsReady & !dadsReady) {
  full_families_moms <- rbind(families, momDF)
  families <- full_families_moms %>% mutate_if(is.numeric, as.integer)
} else if (!momsReady & dadsReady) {
  full_families_dads <- rbind(families, dadDF)
  families <- full_families_dads %>% mutate_if(is.numeric, as.integer)
}
```

We attempt to make a pedigree again:

```{r make ped 2, eval=FALSE}
#bigped <- pedigree(famid = families$kindredID, id = families$egoUPDBID, dad = families$paUPDBID, mom = families$maUPDBID, sex = families$sex, status = families$deceased, affected = families$affection)
```

So it looks like some individuals are missing a parent. We need to create a record for those parents:

```{r id missing one parent}
missing_a_parent <- families %>% filter((paUPDBID != 0 & maUPDBID == 0) | (paUPDBID == 0 & maUPDBID != 0))
both_parents_present <- families %>% filter((paUPDBID == 0 & maUPDBID == 0) | (paUPDBID != 0 & maUPDBID != 0))
```

First, we'll give these people a parent by filling the zeros in `paUPDBID` column with negative numbers:

```{r fill in zeros}
# start here tomorrow, it needs to be do negative numbers on the mothers too
if (0 %in% as.integer(missing_a_parent$paUPDBID)) {
  missing_a_dad <- missing_a_parent %>% filter(paUPDBID == 0)
  missing_a_dad$paUPDBID <- -seq(1:nrow(missing_a_dad))
}
if (0 %in% as.integer(missing_a_parent$maUPDBID)) {
  missing_a_mom <- missing_a_parent %>% filter(maUPDBID == 0)
  missing_a_mom$maUPDBID <- -seq( 1:nrow(missing_a_mom))
}
```

Now we add them back to our dataset:

```{r missing and both}
families <- rbind(both_parents_present, missing_a_dad)
#families <- rbind(families, missing_a_mom)
```

Now we can reuse the above chunks of code to give these 'new' individuals their own records. First, we save what we have so far:

```{r save progress 2}
# if we try to save large numeric types scientific notation is introduced, so we change them to integers
families <- families %>% mutate_if(is.numeric, as.integer)
write_delim(families, "Z:/temp/parentsAdded_medalist_like_ped.tsv", delim = "\t")
```

DUPLICATED CODE BELOW: The following script is used to identify lists of fathers and mothers who do not have their 'own' line (in which their ID is in the `egoUPDBID` column):

```{python add missing parents}
with open('Z:/temp/parentsAdded_medalist_like_ped.tsv') as full_fam_ped:
    full_fam_ped.readline()
    egoDict = {}
    momDict = {}
    dadDict = {}
    for line in full_fam_ped:
        line = line.strip().split("\t")
        family = int(line[0])
        ego = int(line[1])
        dad = int(line[2])
        mom = int(line[3])
        if ego != 0:
            if ego not in egoDict:
                egoDict[ego] = [family]
            else:
                egoDict[ego].append(family)
        if dad != 0:
            if dad not in dadDict:
                dadDict[dad] = [family]
            else:
                dadDict[dad].append(family)
        if mom != 0:
            if mom not in momDict:
                momDict[mom] = [family]
            else:
                momDict[mom].append(family)
    missingDads = []
    missingMoms = []
    for mom, fams in momDict.items():
        for fam in fams:
            if mom in egoDict:
                if fam not in egoDict[mom]:
                    missingMom = [mom, fam]
                    missingMoms.append(missingMom)
            else:
                missingMom = [mom, fam]
                missingMoms.append(missingMom)
        
    for dad, fams in dadDict.items():
        for fam in fams:
            if dad in egoDict:
                if fam not in egoDict[dad]:
                    missingDad = [dad, fam]
                    missingDads.append(missingDad)
            else:
                missingDad = [dad, fam]
                missingDads.append(missingDad)
```

MORE DUPLICATED CODE: We use the lists created by the python script to create new rows. We then append it to our `families` dataframe:

```{r create entries}
missingDads2 <- py$missingDads
missingMoms2 <- py$missingMoms
# no new moms to add!

missingParents <- function(x, y) {
  data.frame(
    kindredID = x[2],
    egoUPDBID = x[1],
    paUPDBID = 0,
    maUPDBID = 0,
    sex = y,
    deceased = 1,
    affection = NA,
    stringsAsFactors = FALSE)
}

#adRows2 <- lapply(missingDads2, missingParents, "male")
#adDF2 <- do.call(rbind, dadRows2)

#23_families_dads2 <- rbind(full_families, dadDF2)

#full_families <- full_families_dads2 %>% mutate_if(is.numeric, as.integer)
if (length(missingDads2) > 0) {
  dadRows2 <- lapply(missingDads2, missingParents, "male")
  dadDF2 <- do.call(rbind, dadRows2)
  dadsReady <- TRUE
} else {
  dadsReady <- FALSE
}

if (length(missingMoms2) > 0) {
  momRows2 <- lapply(missingMoms2, missingParents, "female")
  momDF2 <- do.call(rbind, momRows2)
  momsReady <- TRUE
} else {
  momsReady <- FALSE
}

if (momsReady & dadsReady) {
  full_families_dads2 <- rbind(families, dadDF2)
  full_families_moms_dads2 <- rbind(full_families_dads2, momDF2)
  families <- full_families_moms_dads2 %>% mutate_if(is.numeric, as.integer)
} else if (momsReady & !dadsReady) {
  full_families_moms2 <- rbind(families, momDF2)
  families <- full_families_moms2 %>% mutate_if(is.numeric, as.integer)
} else if (!momsReady & dadsReady) {
  full_families_dads2 <- rbind(families, dadDF2)
  families <- full_families_dads2 %>% mutate_if(is.numeric, as.integer)
}
```

We attempt to make a pedigree object again:

```{r make ped 3}
families <- distinct(families)

families$affection[is.na(families$affection)] <- 0
count <- families %>% group_by(kindredID) %>% summarise(counts = n(), countsAffected = sum(affection))

pedList <- pedigree(famid = families$kindredID, id = families$egoUPDBID, dad = families$paUPDBID, mom = families$maUPDBID, sex = families$sex, status = families$deceased, affected = families$affection)

write_csv(families, "Z:/temp/kinship2_ready_pedigree.csv")
```

It worked! Now we draw the pedigree, inserting the kindredID in the brackets following
pedList

```{r draw pedigrees}

pedigree <- pedList["34479"]

# Does not plot in color
#tiff("Z:/temp/drawn_peds/562503.tiff", units = "in", width = 90, height = 10, res = 300)
#plot.pedigree(pedigree)
#dev.off()

```

Works well up to this point currently...
Add additional info slope, egfr, etc to `fullPed` and plot in color

```{r add additional info to ped}
# load pedigree
preparedPed <- read_csv("Z:/temp/kinship2_ready_pedigree.csv")

# load phenotype data from files. Data can also be read in directly from the database using the RODBC package
allDiags <- read_csv("Y:/IRB_00096551/UDDB_Table_Files_For_Plotting_2021/uddb_diags_2021.csv") %>% separate(UPDB_DIST_ID, c(NA, "UPDB_DIST_ID"), sep = "_")
alleGFRs <- read_csv("Y:/IRB_00096551/UDDB_Table_Files_For_Plotting_2021/uddb_egfrs_2021.csv")
alleGFRs$RESULT_DTM <- as.Date(alleGFRs$RESULT_DTM, format = "%m/%d/%Y")
allSlopes <- read_csv("Y:/IRB_00096551/UDDB_Table_Files_For_Plotting_2021/Patient_with_Slopes_transposed_5-2021.csv", col_types = "cc")
allSlopes$Slope <- as.numeric(allSlopes$Slope)
allACRs <- read_csv("Y:/IRB_00096551/UDDB_Table_Files_For_Plotting_2021/uddb_acrs_2021.csv", col_types = "ccccdc") %>% separate(UPDB_DIST_ID, c(NA, "UPDB_DIST_ID"), sep = "_")
allACRs$RESULT_DTM <- as.Date(allACRs$RESULT_DTM, format = "%m/%d/%Y")
ids <- read_csv("Y:/IRB_00096551/UDDB_Table_Files_For_Plotting_2021/uddb_patientID_2021.csv", col_types = "cnccc") %>% separate(UPDB_ID, c(NA, "UPDB_ID"), sep = "_") %>% separate(DIST_ID, c(NA, "DIST_ID"), sep = "_")
dmTypes <- read_csv("Y:/IRB_00096551/UDDB_Table_Files_For_Plotting_2021/uddb_dmTypes_2021.csv")

# join IDs to the phenotype data
alleGFRs <- left_join(alleGFRs, ids, by = c("Patient_ID"))
allACRs <- left_join(allACRs, ids, by = c("Patient_ID"))
allDiags <- left_join(allDiags, ids, by = c("Patient_ID"))
allSlopes <- left_join(allSlopes, ids, by = c("Patient_ID"))
dmTypes <- left_join(dmTypes, ids, by = c("Patient_ID"))
preparedPed <- left_join(preparedPed, ids[ , c(1,2)], by = c("egoUPDBID" = "PERSON_ID"))

# check if any values are NA. Check how many Patient_IDs contain the string "ID"
sum(grepl("NA", alleGFRs$Patient_ID))
sum(is.na(alleGFRs$Patient_ID))
sum(grepl("NA", allSlopes$Patient_ID))
sum(is.na(allSlopes$Patient_ID))
sum(grepl("NA", allDiags$Patient_ID))
sum(is.na(allDiags$Patient_ID))
sum(grepl("NA", ids$Patient_ID))
sum(is.na(ids$Patient_ID))
sum(grepl("NA", dmTypes$Patient_ID))
sum(is.na(dmTypes$Patient_ID))

# remove the 427 rows with NA in the Patient_ID in Diags
allDiags <- allDiags %>% filter(!is.na(Patient_ID))

# only keep data from the HighRisk family
eGFRHighRisk <- alleGFRs %>% filter(PERSON_ID %in% preparedPed$egoUPDBID)
ACRHighRisk <- allACRs %>% filter(PERSON_ID %in% preparedPed$egoUPDBID)
diagsHighRisk <- allDiags %>% filter(PERSON_ID %in% preparedPed$egoUPDBID)
slopeHighRisk <- allSlopes %>% filter(PERSON_ID %in% preparedPed$egoUPDBID)
dmTypesHighRisk <- dmTypes %>% filter(PERSON_ID %in% preparedPed$egoUPDBID)

# keep all kidney diagnoses from the HighRisk family and find all individuals whose eGFR ever drops below 15
kidneyDiagsHighRisk <- diagsHighRisk %>% filter(ICD_CODE %in% c("585.6", "N18.6", "V42.0", "V45.11", "Z94.0", "Z99.2"))
minEGFRHighRisk <- eGFRHighRisk %>% group_by(Patient_ID, UPDB_ID) %>% summarize(min_egfr = min(EGFR))
observedEGFRHighRisk <- eGFRHighRisk %>% group_by(Patient_ID) %>% summarize(min_egfr = min(EGFR)) %>% filter(min_egfr <= 15)
esrdHighRisk <- unique(c(observedEGFRHighRisk$Patient_ID, kidneyDiagsHighRisk$Patient_ID))

# join slopes to HighRisk
preparedPed <- left_join(preparedPed, slopeHighRisk[ , c(2, 3, 5)], by = c("egoUPDBID" = "PERSON_ID"))

# add other phenotypes to HighRisk
preparedPed$esrd <- ifelse(preparedPed$Patient_ID %in% esrdHighRisk, "ESRD", NA)
#preparedPed$NGS <- ifelse(is.na(preparedPed$NGS), NA, "WGS")
#preparedPed$dmType <- ifelse(preparedPed$egoUPDBID %in% dmTypesHighRisk$PERSON_ID, dmTypesHighRisk$FinalType, NA)
preparedPed <- merge(preparedPed, dmTypesHighRisk, by.x = "egoUPDBID", by.y = "PERSON_ID", all.x = TRUE)
preparedPed$DIST_ID <- preparedPed$DIST_ID.x
preparedPed$Patient_ID <- preparedPed$Patient_ID.x
preparedPed$dmType <- preparedPed$FinalType
preparedPed <- subset(preparedPed, select = -c(FinalType, Patient_ID.x, Patient_ID.y, DIST_ID.x, DIST_ID.y, Type1_count, Type2_count, Type1_Type2_Ratio, UPDB_ID, COHORT_FLAG))

lastegfr <- eGFRHighRisk %>% group_by(Patient_ID) %>% summarize(last_egfr = EGFR[which.max(RESULT_DTM)])
names(lastegfr) <- c("Patient_ID", "last_egfr")
lastacr <- ACRHighRisk %>% group_by(Patient_ID) %>% summarize(last_acr = OBS_VALUE[which.max(RESULT_DTM)])
names(lastacr) <- c("Patient_ID", "last_acr")

preparedPed <- left_join(preparedPed, lastegfr, by = "Patient_ID")
preparedPed <- left_join(preparedPed, lastacr, by = "Patient_ID")

library(kinship2)

preparedPed$rapid_decliner <- ifelse(!is.na(preparedPed$Slope) & preparedPed$Slope <= -5, 1, ifelse(!is.na(preparedPed$esrd), 1, 0))
#preparedPed$affection <- ifelse(preparedPed$affection == 10, 1, 0)
#### I changed this line...
preparedPed <- left_join(preparedPed, minEGFRHighRisk[ , c(1, 3)], by = "Patient_ID")
#preparedPed$affection <- ifelse(is.na(preparedPed$affection), 0, preparedPed$affection)

preparedPed$deceased[is.na(preparedPed$deceased)] <- "0"
preparedPed$deceased[(preparedPed$deceased) == "Y"] <- "1"
#preparedPed$deceased <- ifelse(preparedPed$deceased == "Y", 1, 0)

#pedList <- pedigree(famid = preparedPed$kindredID, id = preparedPed$egoUPDBID, dad = preparedPed$paUPDBID, mom = preparedPed$maUPDBID, sex = preparedPed$sex, status = preparedPed$deceased, affected = preparedPed$rapidDecliner, missid = 0)
pedList <- pedigree(famid = preparedPed$kindredID, id = preparedPed$egoUPDBID, dad = preparedPed$paUPDBID, mom = preparedPed$maUPDBID, sex = preparedPed$sex, status = preparedPed$deceased, affected = preparedPed$rapid_decliner, missid = 0)

##write_tsv(preparedPed, "~/projects/eGFR/HighRisk/HighRisk_files/HighRisk_kinshipReady.tsv")
write_tsv(preparedPed, "Z:/temp/HighRisk_kinshipReady_FINAL.tsv")
```



``` {r final plotting settings}
library(FamAgg)

#library(FamAgg)
#
#newstuff
#ped <- read_tsv("~/temp/HighRisk_kinshipReady_FINAL.tsv")
#preparedPed <- ped[ped$kindredID %in% c(1506, 16366586, 25322898), ]
#preparedPed <- select(preparedPed, c(1:5))
#colnames(preparedPed) <- c("family", "id", "father", "mother", "sex")
#
#preparedPed$og_id <- preparedPed$id
#preparedPed$id <- c(1:nrow(preparedPed))
#uniquePed <- as.data.frame(preparedPed[1:5])
#
##preparedPed$id <- paste(preparedPed$family, preparedPed$id, sep = "_")
#
#fad <- FAData(pedigree = uniquePed)

ped <- read_tsv("Z:/temp/HighRisk_kinshipReady_FINAL.tsv")
#preparedPed <- ped[ped$kindredID == 34478]
preparedPed <- filter(ped, kindredID == 92898)
#preparedPed <- select(preparedPed, c(1:5))
colnames(preparedPed)[1:5] <- c("id", "family", "father", "mother", "sex")
preparedPed <- select(preparedPed, family, id, father, mother, sex, deceased, affection, Slope, esrd, DIST_ID, Patient_ID, dmType, last_egfr, last_acr, rapid_decliner, min_egfr)

#preparedPed$id[preparedPed$id == preparedPed$family] = paste0(preparedPed$id, "_1")
#preparedPed$father[preparedPed$father == preparedPed$family] = paste0(preparedPed$father, "_1")
#preparedPed$mother[preparedPed$mother == preparedPed$family] = paste0(preparedPed$mother, "_1")

#preparedPed$id <- paste(preparedPed$family, preparedPed$id, sep = "_")

#fad <- FAData(pedigree = preparedPed)

#dups <- group_by(preparedPed, id) %>% summarise(count = n())
#kindred <- group_by(ped, kindredID) %>% summarise(count = n())

x <- "92898"
partialPed <- filter(fullPed, kindredID == as.integer(x))
fullInfo <- merge(preparedPed, partialPed, by.x = "id", by.y = "egoUPDBID", all.x = TRUE)

# Pull the columns we will use later and prep to merge with Byr
fullInfoFirstCols <- fullInfo[1:16]
fullInfoFirstCols <- select(fullInfoFirstCols, family, id, father, mother, sex.x, deceased, affection.x, Slope, esrd, DIST_ID.x, Patient_ID, dmType, last_egfr, last_acr, rapid_decliner, min_egfr)
fullInfoFirstCols$Byr <- fullInfo$Byr
colnames(fullInfoFirstCols)[5] = "sex"
colnames(fullInfoFirstCols)[7] = "affection"
colnames(fullInfoFirstCols)[10] = "DIST_ID"

# Calculate current age using current date...
fullInfo <- fullInfoFirstCols
fullInfo$year <- 2021
fullInfo$age <- ifelse(fullInfo$deceased == 0, fullInfo$year - fullInfo$Byr, "deceased")

# Remove Byr and year unnecessary columns
fullInfo <- subset(fullInfo, select = -c(Byr, year))

fullInfo$color <- ifelse(is.na(fullInfo$dmType), 1,
                         ifelse(grepl("T2D", fullInfo$dmType) | grepl("T3D", fullInfo$dmType), 2,
                                ifelse(grepl("T1D", fullInfo$dmType), 4, 5)))
                         
pedigree <- pedList[as.character(x)]
pdf(paste0("Z:/temp/drawn_peds/annotated_", x, "_high_risk.pdf"), width = 60, height = 12)
#pdf(paste0("~/projects/eGFR/HighRisk/annotated_", x, "_fras1.pdf"), width = 24, height = 12)

# this might need to change
#listOfIds <- ifelse(is.na(fullInfo$Slope) & is.na(fullInfo$dmType) & fullInfo$Study_ID != "FS180003", 
listOfIds <- ifelse(is.na(fullInfo$dmType), 
                  paste0(fullInfo$id, "\n", fullInfo$age), 
                  paste0(fullInfo$id, "\n",fullInfo$dmType, "\n", fullInfo$min_egfr, "\n", fullInfo$last_egfr, "\n",
                         fullInfo$Slope, "\n", fullInfo$esrd, "\n", fullInfo$age))
                  #paste0(fullInfo$UPDB_ID, "\n", fullInfo$Dist_ID, "\n", fullInfo$Study_ID, "\n", 
                  #       fullInfo$dmType, "\n", fullInfo$min_egfr, "\n", fullInfo$last_egfr, "\n", 
                  #       fullInfo$last_acr, "\n", fullInfo$Slope, "\n", fullInfo$esrd, "\n", fullInfo$NGS))

plot.pedigree(pedigree, id = listOfIds, col = fullInfo$color, affected = fullInfo$affection, cex = 0.5, symbolsize = 1.75)

mtext(paste0("Family ", x), line = 2, adj = 0, cex = 2.5)

dev.off()
```